Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-23 paper/sec-mva.tex
updated plots

From Table~\ref{table:cutflowMVA} we see that
after the application of the MVA, 
the signal significance in the boosted category increases
from 1.2 to 5.4, with $S/B$ increasing from $0.14\%$ to $3.4\%$,
with almost 800 signal events expected at the HL-LHC.
%
For the intermediate and resolved categories, $S/\sqrt{B}$
increases from 0.9 and 1.0 to 4.8 and 4.5, with
the signal over background ratio raising from
$0.12\%$ and $0.03\%$ to 6\% and 1\%, respectively.
%
The overall combined significance is thus $S\sqrt{B}\simeq 8$,
well above the threshold for discovery. However such a significance
is unrealistic given that the HL-LHC will be a high-PU environment,
which will affect the description of the various
kinematical distributions used as input to the MVA.
%
Therefore we shall now quantify the robustness of the
results summarized Table~\ref{fig:sb_mva}
in a realistic environment including the effects of PU.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{|c|l|c|c|c|c|}
        \hline
     \multicolumn{6}{|c|}{$\la n_{\rm PU}\ra=80$+SK} \\
     \hline
         \hline
    Category  &   &  $N_{\rm ev}$ signal &  $N_{\rm ev}$ back  &  $S/\sqrt{B}$ & $S/B$ \\ 
    \hline
    \hline
    \multirow{2}{*}{Boosted} &  $y_{\rm cut}=0$  & 360   &  $1.4\cdot 10^6$ & 0.31   &
     $3\cdot 10^{-4}$  \\
    &  $y_{\rm cut}=0.82$ &  250 & 7300  & 2.9    & 0.034  \\
    \hline
    \hline
    \multirow{2}{*}{Intermediate} &  $y_{\rm cut}=0$  &  150  & $9\cdot 10^5$    & 0.15    &
     $2\cdot 10^{-4}$ \\
    &  $y_{\rm cut}=0.80$ & 50 & 2600  &  1.1   & 0.02 \\
    \hline
    \hline
    \multirow{2}{*}{Resolved} &  $y_{\rm cut}=0$  &  1100  & $8.5\cdot 10^6$
    & 0.25    &  $1.3\cdot 10^{-4}$  \\
    &  $y_{\rm cut}=0.65$ & 430  & $3\cdot 10^4$  &  2.5   & 0.015  \\
    \hline
      \end{tabular}
  \caption{\small Same as Table~\ref{table:cutflowMVA}, now for the case
    of PU+SK, with $\la n_{\rm PU}\ra=80$.
        \label{table:cutflowMVA_PU}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.49\textwidth]{plots/roc_SKPU80.pdf}
\includegraphics[width=0.49\textwidth]{plots/nev2_SKPU80.pdf}
\caption{\small Same as Fig.~\ref{fig:nev2} in the
case of events with PU, for
$\la n_{\rm PU}\ra=80$ using
{\tt SoftKiller} for PU subtraction.
}
\label{fig:nev2_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.48\textwidth]{plots/ssb_SKPU80.pdf}
\includegraphics[width=0.48\textwidth]{plots/sb_SKPU80.pdf}
\caption{\small 
Same as Fig.~\ref{fig:sb_mva} in the
case of events with PU, for
$\la n_{\rm PU}\ra=80$
using
{\tt SoftKiller} for PU subtraction.
}
\label{fig:sb_mva_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.49\textwidth]{plots/res_wgthist_SKPU80.pdf}
\includegraphics[width=0.49\textwidth]{plots/bst_wgthist_SKPU80.pdf}
\vspace{-0.5cm}
\caption{\small
Same as Fig.~\ref{fig:nnweights} in the
case of events with PU, for
$\la n_{\rm PU}\ra=80$
using
{\tt SoftKiller} for PU subtraction.
}
\label{fig:nnweights_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
  \centering
  \small
  \begin{tabular}{|c|c|c|c||c|c||c|c|}
        \hline
        Category  &    &  $N_{\rm ev}^{\rm tot}$  &  $N_{\rm ev}^{\rm 4b}$   &
        $S/\sqrt{B_{\rm tot}}$ & $S/\sqrt{B_{\rm 4b}}$  
        &  $S/B_{\rm tot}$ & $S/B_{\rm 4b}$\\ 
    \hline
    \hline
    \multirow{2}{*}{Boosted} &  no PU  & $2.2\cdot 10^4$  & $1.5\cdot 10^4$     & 
      5.4 &  6.5 & 0.034 & 0.05 \\
    & PU+SK & 7300  &  500  &  2.9  & 3.6 &  0.034 & 0.04 \\
    \hline
    \hline
    \multirow{2}{*}{Intermediate} &  no PU   & 4500  & 2000    &
    4.8  &  5.3 &  0.06  &  0.08 \\
    & PU+SK  & 2600   &  1500  & 1.1  & 1.2 & 0.02 & 0.03 \\
    \hline
    \hline
    \multirow{2}{*}{Resolved} &   no PU  & $8\cdot 10^4$   &
    $4.5\cdot 10^4$
    & 4.5  & 6.6  & 0.01 & 0.03 \\
    & PU+SK  &  $3\cdot 10^4$   &   $1.3\cdot 10^4$ & 2.5    & 3.7  & 0.015 & 0.03 \\
    \hline
      \end{tabular}
  \caption{\small Post-MVA number of background events
    expected at the HL-LHC, comparing the total number, $N_{\rm ev}^{\rm tot}$,
    with those from the QCD $4b$ background
    component only, $N_{\rm ev}^{\rm 4b}$.
     %
    Also provided are the values of the signal 
    significance and signal over background ratios in each case.
        \label{table:cutflowMVA_fakes}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Combining the three categories, taking into
account all background components, we obtain the overall signal
significance:
\be
\lp \frac{S}{\sqrt{B}}\rp_{\rm tot} \simeq 4.0~(1.3) \, ,\quad
\mathcal{L}=3000~(300)\,{\rm fb}^{-1}\, ,
\ee
%
indicating  that the measurement of
Higgs pair production in the $b\bar{b}b\bar{b}$ final state at the HL-LHC
should be 
well above the threshold for evidence.
%
On the other hand, such a measurement seems to be
extremely challenging
for the total integrated luminosity expected by the end of Run II.
%
Under the  assumption that
    the only relevant background would be the irreducible QCD $4b$ component,
    we obtain instead:
    \be
\lp \frac{S}{\sqrt{B_{\rm 4b}}}\rp_{\rm tot} \simeq 5.3~(1.8) \, ,\quad
\mathcal{L}=3000~(300)\,{\rm fb}^{-1}\, ,
\ee
Therefore, a discovery of Higgs pair production at the
HL-LHC might be possible if the
mis-identification of light and charm jets as
$b$-jets can be  reduced.
%
With these assumptions, and provided
the signal efficiency can be further enhanced,
this measurement might even be possible
by the end of Run II.
%
The smaller value of $\la n_{\rm PU}\ra$ at Run II as compared
to the HL-LHC enviroment should be beneficial in this
respect.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

From Table~\ref{table:cutflowMVA} we see that
after the application of the MVA, 
the signal significance in the boosted category increases
from 1.2 to 5.4, with $S/B$ increasing from $0.14\%$ to $3.4\%$,
with almost 800 signal events expected at the HL-LHC.
%
For the intermediate and resolved categories, $S/\sqrt{B}$
increases from 0.9 and 1.0 to 4.8 and 4.5, with
the signal over background ratio raising from
$0.12\%$ and $0.03\%$ to 6\% and 1\%, respectively.
%
The overall combined significance is thus $S\sqrt{B}\simeq 8$,
well above the threshold for discovery.
%
However, given that the HL-LHC will be a high-PU environment,
which will affect the description of the various
kinematical distributions used as input to the MVA,
it is essential to quantify the robustness of these
results
in a realistic environment including the effects of PU.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{|c|l|c|c|c|c|}
        \hline
     \multicolumn{6}{|c|}{$\la n_{\rm PU}\ra=80$+SK+Trim} \\
     \hline
         \hline
    Category  &   &  $N_{\rm ev}$ signal &  $N_{\rm ev}$ back  &  $S/\sqrt{B}$ & $S/B$ \\ 
    \hline
    \hline
    \multirow{2}{*}{Boosted} &  $y_{\rm cut}=0$  & 1000   &  $4.5\cdot 10^5$ & 1.5   & $2\cdot 10^{-3}$  \\
    &  $y_{\rm cut}=0.7$ &  780  & $4.5\cdot 10^4$  & 3.7    & 0.02  \\
    \hline
    \hline
    \multirow{2}{*}{Intermediate} &  $y_{\rm cut}=0$  &  630  & $8.2\cdot 10^5$    & 0.7    &
     $8\cdot 10^{-4}$ \\
    &  $y_{\rm cut}=0.7$ & 380 & $8.5\cdot 10^3$  &  4.1   & 0.04 \\
    \hline
    \hline
    \multirow{2}{*}{Resolved} &  $y_{\rm cut}=0$  &  4500  & $2.7\cdot 10^7$
    & 0.9    &  $2\cdot 10^{-4}$  \\
    &  $y_{\rm cut}=0.5$ & 2000  & $2.1\cdot 10^5$  &  4.6   & 0.01  \\
    \hline
      \end{tabular}
  \caption{\small Same as Table~\ref{table:cutflowMVA}, now for the case
    of PU+SK+Trim, with $\la n_{\rm PU}\ra=80$.
        \label{table:cutflowMVA_PU}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.49\textwidth]{plots/roc_SKPU80.pdf}
\includegraphics[width=0.49\textwidth]{plots/nev2_SKPU80.pdf}
\caption{\small Same as Fig.~\ref{fig:nev2} in the
case of events with PU, for
$\la n_{\rm PU}\ra=80$ with SK+Trim.
}
\label{fig:nev2_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.48\textwidth]{plots/ssb_SKPU80.pdf}
\includegraphics[width=0.48\textwidth]{plots/sb_SKPU80.pdf}
\caption{\small 
Same as Fig.~\ref{fig:sb_mva} in the
case of events with PU, for
$\la n_{\rm PU}\ra=80$ with SK+Trim.
}
\label{fig:sb_mva_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.49\textwidth]{plots/res_wgthist_SKPU80.pdf}
\includegraphics[width=0.49\textwidth]{plots/bst_wgthist_SKPU80.pdf}
\vspace{-0.5cm}
\caption{\small
Same as Fig.~\ref{fig:nnweights} in the
case of events with PU, for
$\la n_{\rm PU}\ra=80$ with SK+Trim.
}
\label{fig:nnweights_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
  \centering
  \small
  \begin{tabular}{|c|c|c|c||c|c||c|c|}
        \hline
        Category  &    &  $N_{\rm ev}^{\rm tot}$  &  $N_{\rm ev}^{\rm 4b}$   &
        $S/\sqrt{B_{\rm tot}}$ & $S/\sqrt{B_{\rm 4b}}$  
        &  $S/B_{\rm tot}$ & $S/B_{\rm 4b}$\\ 
    \hline
    \hline
    \multirow{2}{*}{Boosted} &  no PU  & $2.2\cdot 10^4$  & $1.5\cdot 10^4$     & 
      5.4 &  6.5 & 0.03 & 0.05 \\
    & PU+SK &  $4.5\cdot 10^4$ &  $1.7\cdot 10^4$    &  3.7  & 5.9 &  0.02 & 0.04 \\
    \hline
    \hline
    \multirow{2}{*}{Intermediate} &  no PU   & $4.5\cdot 10^3$  & $2.0\cdot 10^3$    &
    4.8  &  5.3 &  0.06  &  0.08 \\
    & PU+SK+Trim  & $8.5\cdot 10^3$   &  $4.6\cdot 10^3$  & 4.1  & 5.6 & 0.04 & 0.08 \\
    \hline
    \hline
    \multirow{2}{*}{Resolved} &   no PU  & $8.1\cdot 10^4$   &
    $4.5\cdot 10^4$
    & 4.5  & 6.6  & 0.01 & 0.03 \\
    & PU+SK+Trim  &  $2.1\cdot 10^5$   &   $1.1\cdot 10^5$ & 4.6    & 6.2  & 0.01 & 0.02 \\
    \hline
      \end{tabular}
  \caption{\small Post-MVA number of background events
    expected at the HL-LHC, comparing the total number, $N_{\rm ev}^{\rm tot}$,
    with those from the QCD $4b$ background
    component only, $N_{\rm ev}^{\rm 4b}$.
     %
    Also provided are the values of the signal 
    significance and signal over background ratios in each case.
        \label{table:cutflowMVA_fakes}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Combining the three categories, taking into
account all background components, we obtain the overall signal
significance:
\be
\lp \frac{S}{\sqrt{B}}\rp_{\rm tot} \simeq 7.2~(2.3) \, ,\quad
\mathcal{L}=3000~(300)\,{\rm fb}^{-1}\, ,
\ee
%
indicating  that the measurement of
Higgs pair production in the $b\bar{b}b\bar{b}$ final state at the HL-LHC
should be 
well above the threshold for evidence.
%
On the other hand, such a measurement seems to be
extremely challenging
for the total integrated luminosity expected by the end of Run II.
%
Under the  assumption that
    the only relevant background would be the irreducible QCD $4b$ component,
    we obtain instead:
    \be
\lp \frac{S}{\sqrt{B_{\rm 4b}}}\rp_{\rm tot} \simeq 10.2~(3.2) \, ,\quad
\mathcal{L}=3000~(300)\,{\rm fb}^{-1}\, ,
\ee
Therefore, a discovery of Higgs pair production at the
HL-LHC might be possible if the
mis-identification of light and charm jets as
$b$-jets can be  reduced.
%
With these assumptions, and provided
the signal efficiency can be further enhanced,
this measurement might even be possible
by the end of Run II.
%
The smaller value of $\la n_{\rm PU}\ra$ at Run II as compared
to the HL-LHC environment should be beneficial in this
respect.
